Report

### 2022-06-27 manuscript.Rmd
expand tutorial

### The Bayes Factor using bain.
In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis. This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist. However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much. In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis. Because the focus of this tutorial is on BES, this tutorial will only briefly introduce how to evaluate an hypothesis using the R package `bain` and interpret its output.  For an extensive tutorial on how to evaluate informative hypotheses using `bain` see [@hoijtink2019tutorial].

Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the following null-hypothesis:

@@

However, we may expect that the children growing up in an advantaged site will have a higher score than children growing up in a disdvantaged site. Our informative hypothesis is:

$H_A: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$

@@

fit_bain <- bain(fit, hypothesis = "site1=site2=site3=site4=site5; site2&site3>site1&site4&site5" )
print(fit_bain)

```
<!-- Both BF.c and BF.u are close to zero, the probability of the HA being more probable than the Hc or Hu is small. <further interpretation>-->

### Single vs multiple studies.
In the case of multiple studies, it is possible to evaluate the hypotheses under consideration by synthesizing the evidence over all studies. In the example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same subject.
In this example, each result has the $H_0$ and $H_A$ in common, but contains one result-specific hypothesis. The `pbf()` function only gives output for the hypotheses that are present in all results.
```{r}
anov <- lm(postnumb~site-1,sesamesim[1:75,])
results <- bain(anov, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
anov2 <- lm(postnumb~site-1,sesamesim[76:150,])
results2 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
anov3 <- lm(postnumb~site-1,sesamesim[151:nrow(sesamesim),])
results3 <- bain(anov, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")

pbf(list(results, results2, results3))

```
The function returns the individual BF.c's of each study, as well as their product.

>>>>

### The Bayes Factor using bain.
In conventional Null-hypothesis significance testing (NHST), as the name implies, data is analysed to generate evidence for or against the null-hypothesis. This method can only evaluate a single hypothesis, the null-hypothesis, and its output is dichotomous in nature: whether a theoretical relationship is probable or improbable to exist given the data. However, there are many more hypotheses that can theoretically be tested and researchers may not be interested in the null-hypothesis. They may also not only be interested in whether the hypothesis is probable, but also by how much. In the Bayesian framework, the Bayes Factor (BF) is a test statistic that gives the amount of support for any possible desired hypothesis. Because the focus of this tutorial is on BES, this tutorial will only briefly introduce how to evaluate an hypothesis using the R package `bain` and interpret its output. For an extensive tutorial on how to evaluate informative hypotheses using `bain` see [@hoijtink2019tutorial].

Let's say we are interested in whether the status of the site children grew up in (`site`) affects their score on a numbers test after watching Sesame Street for a year (`postnumb`). `postnumb` is a continuous variable which ranges from 0-63. `site` is a nominal variable indicating group membership with possible values 1 = disadvantaged inner city, 2 = advantaged suburban, 3 = advantaged rural, 4 = disadvantaged rural and 5 = disadvantaged Spanish speaking. In NHST we specify the null-hypothesis that the mean `postnumb` is equal for all groups in  site`:

@@

However, we may have prior expectations that the children growing up in an advantaged site will have a higher score than children growing up in a disadvantaged site. Our informative hypothesis is:

$H_i: \mu_{site2},\mu_{site3} > \mu_{site1},\mu_{site4},\mu_{site5}$

@@

#specify Hypotheses to evaluate
hypotheses <- "site2=site1=site3=site4=site5; site2&site3>site1&site4&site5"
hypotheses <- "site2=site1=site3=site4=site5"

fit_bain <- bain(fit, hypothesis = hypotheses)
print(fit_bain)

```
The output of `bain` seems to support neither hypothesis. The `fit` for both hypotheses is approximately zero which means that almost none of the data are in agreement with the restrictions the specified hypotheses pose. The `BF.c`, which is the Bayes Factor of the hypothesis versus its complement, is for both hypotheses close to zero. This implies that the complement of both hypotheses is $\frac{1}{BF.c}$ times more likely than the passed hypotheses. Finally, `PMPc`, the Posterior Model Probabilities of the specified hypotheses with $H_c$ included, is also close to zero for both specified hypothesis. This means that *out of* the specified hypotheses and their complement, the probability of $H_c$ being true is 99.9%.

### Single vs multiple studies.
Sometimes there is access to the data of not only a single, but multiple studies that all contain the necessary information to evaluate our hypothesis $H_i$. To evaluate $H_i$, it is possible to synthesize the support for $H_i$ over all studies using the PBF. In this example we split the `sesamesim` dataset up into three equal parts, to simulate three studies into the same topic.

```{r}
sample1 <- sesamesim[1:75,]
sample2 <- sesamesim[76:150,]
sample3 <- sesamesim[151:nrow(sesamesim),]

In the first use-case of `pbf()`, an `lm()` is ran for each sample. All `lm` objects are stored in a list which is passed to `pbf()`.
`pbf()` supports lists of all objects that `bain()` also supports. All supported types are specified in bain vignette *REF*. One exception is the use of named vectors, which will be covered later.
```{r}
lm_sample1 <- lm(postnumb~site-1, sample1)
lm_sample2 <- lm(postnumb~site-1, sample2)
lm_sample3 <- lm(postnumb~site-1, sample3)

lm_allsamples <- list(lm_sample1, lm_sample2, lm_sample3)

pbf(lm_allsamples, hypotheses)
```

The function returns a `data.frame` containing the passed hypotheses in the rows and in the columns the PBF and the individual Bayes Factor for each sample individually. The `PBF` of both hypotheses is close to zero, meaning that the combined evidence for both hypotheses over all studies is low. When looking at the individual Bayes Factors, we also see that the PBF is lower than all of the Bayes Factors individually. For sample 3 the specified hypotheses are even more likely than their complement. This example illustrates another strength of the PBF; being able to evaluate a hypothesis more extremely than the most extreme individual evaluation. If all samples do not support a hypothesis, than the number of times the hypothesis is not supported should be included in its evaluation. This leads to a more extreme combined evaluation than any individual evaluation. Likewise, if there is very strong support in only a small minority of the samples, while the majority of samples shows no support at all, than it is likely that the hypothesis is not true.

### using bain objects
The `pbf()` function also works when objects are first converted to `bain` objects. In this specific example, each `bain` object corresponding to one of the three samples has the $H_0$ and $H_A$ in common, but contains one sample-specific hypothesis. The `pbf()` function only returns output for the hypotheses that are present in all results. When `bain` objects are passed to the `pbf()`, no additional hypotheses need to be passed to `pbf()` as they are already saved in the objects.

```{r}
bain_sample1 <- bain(lm_sample1, "site2=site1=site3=site4=site5; site2>site1=site3=site4=site5; site2&site3>site1&site4&site5")
bain_sample2 <- bain(lm_sample2, "site2=site1=site3=site4=site5; site2=site1=site3>site4=site5; site2&site3>site1&site4&site5")
bain_sample3 <- bain(lm_sample3, "site2=site1=site3=site4=site5; site2=site1=site3=site4>site5; site2&site3>site1&site4&site5")

bain_allsamples <- list(bain_sample1, bain_sample2, bain_sample3)

pbf(bain_allsamples)

As can be seen, the results are identical to the results in the previous example. This means that the sample-specific hypotheses have been left out and only the hypotheses that each sample has in common are retained.

When there is no hypothesis that is shared by each sample, `pbf()` throws an error.
```{r}
bains_different_hyps <- list(
  bain(lm_sample1, "site2=site1=site3=site4=site5"),
  bain(lm_sample2, "site2=site1=site3>site4=site5"),
  bain(lm_sample3, "site2&site3>site1&site4&site5")
)

tryCatch({
  pbf(bains_different_hyps)},
  error = function(e){message(e)}
      )

```
The function returns an error stating that the samples have no hypothesis in common and thus the PBF cannot be calculated.

### using named vectors
Another specific use-case is when the raw data from different samples is not available. In this case, the researcher only has access to the parameter-values that are provided in the research papers. Because the `pbf()` function inherently converts objects to `bain` objects, each sample requires three pieces of information in order to be usable: the named parameter estimates, a covariance matrix of their standard errors and the sample size. First these values are extracted from all samples. A function is created that extracts is information which is then used to extract the necessary information from all samples.

```{r}
get_estimates <- function(sample){
  number_sites <- length(levels(sesamesim$site))

  sample_size <- nrow(sample)

  site_estimates <- sapply(1:number_sites, function(i){
    rbind(
      mean(sample[sample['site'] == i, "postnumb"]),
      sd(sample[sample['site'] == i, "postnumb"])
      )
    })

  est_names <- sprintf("site%d", 1:number_sites)
  site_means <- site_estimates[1,]
  names(site_means) <- est_names

  covmat <- diag((site_estimates[2,] / sqrt(sample_size))^2)
  dimnames(covmat) <- list(est_names, est_names)
  return(list(means = site_means, covmat = covmat, sample_size = sample_size))
}

ests_sample1 <- get_estimates(sample1)
ests_sample2 <- get_estimates(sample2)
ests_sample3 <- get_estimates(sample3)

Then, each object is converted to a `bain` object with these estimates
```{r}
bain_sample1_ests <- bain(x = ests_sample1[['means']], 
                          Sigma = ests_sample1[['covmat']], 
                          n = ests_sample1[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample2_ests <- bain(x = ests_sample2[['means']], 
                          Sigma = ests_sample2[['covmat']], 
                          n = ests_sample2[['sample_size']], 
                          hypothesis = hypotheses)

bain_sample3_ests <- bain(x = ests_sample3[['means']], 
                          Sigma = ests_sample3[['covmat']], 
                          n = ests_sample3[['sample_size']], 
                          hypothesis = hypotheses)

bain_allsamples_ests = list(bain_sample1_ests, bain_sample2_ests, bain_sample3_ests)
```

And finally, on these objects, `pbf()` can be called.
```{r}
pbf(bain_allsamples_ests)
```
<!-- It is interesting that these estimates are not identical to the other examples... Is that because we do not have ALL information available in this use-case, or is something off in the script? -->
